May I Ask Where You Have Been, Photon?

### 2021-11-20 article.tex
almost finalise article, references

\usepackage{svg,subfig,framed}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

In classical physics, a particle travelling between two points has a single
path, which can be calculated by specifying its state at either endpoint. On 
the other hand, in quantum physics, the trajectory a particle takes between two 
measurements is a weighted sum of every path it could have taken between the 
two measurements. This gives rise to the phenomenon of a particle interfering 
with itself before being measured, as demonstrated by experiments like the 
single-particle version of Young's double slit experiment.

First performed in the early 19th century, the double experiment involves a 
relatively simple procedure: a beam of light is passed through two closely 
spaced slits and projected onto a screen, and the form of the light incident on 
the screen is observed. Because of the wave nature of light, different 
distances travelled by different wavefronts give rise to phase differences at 
the screen, causing an interference pattern to appear, as illustrated in 
figure~\ref{fig:uncollimated}. This appearance of light and dark bands is 
similar to how ripples interfere in a pond, creating alternating zones of 
maximal and minimal displacement based on how crests and troughs overlap.

@@

\A Because the particle underwent wave-like interference with itself, it can be 
	said that it took a "superposition" of \textit{all possible paths} 
between the light source and the screen, meaning that it travelled through both 
the slits. \footnotemark \end{framed}

@@

The Tel-Aviv experiment calls even this "common sense" assertion into question.  
The authors use the Mach-Zehnder interferometer, a device similar to Young's 
apparatus but better suited to precise theoretical treatment in the framework 
of quantum mechanics. It consists of a light source and a detector along with 
two beam splitters (basically half-silvered mirrors) and two mirrors, arranged 
as shown in figure~\ref{fig:interferometer1}. In the classical version of this 
experiment, the light source emits a continuous beam of light, which partly 
passes through and is partially reflected by the first beam splitter. The two 
beams of half the original intensity that emerge are then directed by the two 
mirrors to meet at the second beam splitter. Through a purely classical 
phenomenon, reflection at the vacuum-silver interface of a mirror causes a 
phase change to take place, causing the beams coming out of the second beam 
splitter to combine totally constructively on one side, and totally 
destructively on the other, in analogy to the bright and dark bands in the 
classical double slit experiment.

Extending the analogy to the double-slit experiment, if the light source in the 
Mach-Zehnder interferometer is tuned to a low-enough intensity that single 
photons emitted can be resolved by the detector, it is found that the photon 
arrives at the detector 100\% of the time, demonstrating the phenomenon of a 
particle interfering with itself.  Of course, if we remove the second beam 
splitter, there is a 50\% chance that the photon will show up in our detector, 
and a 50\% chance that it will come out of the other branch of the 
interferometer.

Again, we can answer titular question using our "common sense" approach:

\begin{framed}
\Q May I ask where you have been, photon?

\A The photon's trajectory in the interferometer is a superposition of the 
	classical trajectories through the upper and lower arms of the 
	interferometer.
\end{framed}

@@

Danan et al. start off with a slightly modified version of the Mach-Zehnder 
interferometer, as shown in figure~\ref{fig:danan-1a}. To mark each possible 
path of the photon, they make each mirror oscillate vertically with a unique 
frequency.  Then, to make it possible to monitor the frequency with which the 
detected light oscillates, they use a stacked array of detectors in place of a 
simple light detector. As expected, when the experiment is performed with these 
simple alterations, the power spectrum shows equal contributions at frequencies 
$f_A$ and $f_B$ associated with mirrors $A$ and $B$. Whilst this does not 
directly indicated where individual photons have been---the photon source emits 
numerous photons, meaning that different photons could have caused the peaks at 
$f_A$ and $f_B$---it at least confirms that some of the photons have been at 
each mirror.

@@

The surprising result from this experiment is obtained when the following 
changes are made, as illustrated in figure \ref{fig:danan-2b}:
either the second beam splitter of the miniature Mach-Zehnder interferometer is 
flipped, or mirror $B$ is moved to slightly adjust the path length, causing 
destructive interference of the light propagating towards $F$.

@@

>>>>

\usepackage{svg,subfig,framed,braket}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

In classical physics, a particle travelling between two points has a single
trajectory, which can be calculated by specifying its state at either endpoint.  
On the other hand, in quantum physics, performing a measurement on a particle 
at the beginning and the end of a period (so-called \textit{pre-} and 
\textit{post-selection}) only restricts the the trajectory a particle takes to 
a weighted sum of every path it could have taken.\footnote{It is significant 
that this procedure is not possible in the classical world---there, measuring a 
particle's state at one endpoint gives us information about its whole 
trajectory} This gives rise to the phenomenon of a particle interfering with 
itself before being measured, as demonstrated by experiments like the 
single-particle version of Young's double slit experiment.

First performed in the early 19th century, the double slit experiment involves 
a relatively simple procedure: a beam of light is passed through two closely 
spaced slits and projected onto a screen, and the form of the light incident on 
the screen is observed. Because of the wave nature of light, different 
distances travelled by different wavefronts give rise to phase differences at 
the screen, causing an interference pattern to appear, as illustrated in 
figure~\ref{fig:uncollimated}. This appearance of light and dark bands is 
similar to how ripples interfere in a pond, creating alternating zones of 
maximal and minimal displacement based on how the crests and troughs overlap.

@@

\A Because the particle underwent wave-like interference with itself, it can be 
	said that it took a \textit{superposition} of \textit{all possible 
paths} between the light source and the screen, meaning that it travelled 
through \textit{both} slits.  \footnotemark \end{framed}

@@

The Tel-Aviv experiment calls even this "common sense" reasoning into question.  
The authors use the Mach-Zehnder interferometer, a device similar to Young's 
apparatus but better suited to precise theoretical treatment in the framework 
of quantum mechanics. It consists of a light source and a detector along with 
two beam splitters (basically half-silvered mirrors) and two mirrors, arranged 
as shown in figure~\ref{fig:interferometer1}. In the classical version of this 
experiment, the light source emits a continuous beam of light, which partly 
passes through and is partly reflected by the first beam splitter. The two 
beams (of half the original intensity) that emerge are then directed by the two 
mirrors to meet at the second beam splitter. Through a purely classical 
phenomenon, reflection at the vacuum-silver interface of a mirror causes the 
light to change phase (see labels on figure~\ref{fig:interferometer1}). A 
sequence of these phase changes causes the beams coming out of the second beam 
splitter to combine totally constructively on one side, and totally 
destructively on the other, analogous to the bright and dark bands in the 
classical double slit experiment.

Extending this analogy, if the light source in the Mach-Zehnder interferometer 
is tuned to a low-enough intensity that single photons emitted can be resolved 
by the detector, it is found that the photon arrives at the detector 100\% of 
the time, demonstrating the phenomenon of a particle interfering with itself.  
Of course, if we remove the second beam splitter, there is a 50\% chance that 
the photon will show up in our detector, and a 50\% chance that it will come 
out of the other branch of the interferometer. Again, we can attempt to answer 
the titular question using our "common sense" approach:

\begin{framed}

	\Q Where have you been, photon?

	\A Like in the double-slit experiment, each photon undergoes wave-like 
	interference with itself. The trajectory of the photon between the 
	light source and the light detector is a superposition of a trajectory 
	through each arm of the interferometer.

\end{framed}

@@

Danan et al. start off with a slightly modified version of the Mach-Zehnder 
interferometer, as shown in figure~\ref{fig:danan-1a}. To mark each possible 
path of the photon, they make each mirror oscillate vertically with a unique 
frequency. Then, to make it possible to monitor the frequency with which the 
detected light oscillates (so called "weak measurement"), they use a stacked 
array of detectors in place of a simple light detector. As expected, when the 
experiment is performed with these simple alterations, the power spectrum shows 
equal contributions at frequencies $f_A$ and $f_B$ associated with mirrors $A$ 
and $B$. Whilst this does not directly indicated where individual photons have 
been---the photon source emits numerous photons, meaning that different photons 
could have caused the peaks at $f_A$ and $f_B$---it at least confirms that some 
of the photons have been at each mirror.

@@

@@

The surprising result from this experiment is obtained when the following 
changes are made, as illustrated in figure \ref{fig:danan-2b}:
either the second beam splitter of the miniature Mach-Zehnder interferometer is 
flipped, or mirror $B$ is moved to slightly adjust the path length, causing 
destructive interference of the light propagating towards $F$. Seemingly, the 
only possible path the light could take to reach the detector is via mirror 
$C$. Yet, the observed spectrum contains frequency signatures from mirrors $A$ 
and $B$. To verify the destructive interference of light directed towards $F$, 
the path from $C$ to the detector is blocked (figure~\ref{fig:danan-2c}), 
yielding a null result. Somehow, light from the nested Mach-Zehnder 
interferometer is making its way to the detector! Furthermore, even though the 
photon somehow "visits" mirrors $A$ and $B$, it avoids picking up the frequency 
signatures of mirrors $E$ and $F$! How?

\begin{framed}

	\Q Where have you been, photon?

	\A Now, it becomes hard to say. Using the same reasoning with which we 
	answered our question for the double slit experiment, it seems that the 
	light should be coming from the upper branch of the interferometer.  
	Consistent with this, upon blocking light emerging from mirror $C$, the 
	signal in the upper branch disappears. Yet, upon allowing light to 
	propagate through the upper branch, the frequencies $f_A$ and $f_B$ are 
	unexpectedly present in the power spectrum.

\end{framed}

There have been arguments against this experimental procedure, claiming that 
the vibration of the mirrors prevents full quantum interference from occuring.  
However, Danan et al. argue that a simpler explanation is possible, using 
description of quantum mechanics called the Two-State Vector Formalism (TSVF).

In the standard quantum mechanics formalism that undergraduate students of 
physics are familiar with, if a quantum system is measured (pre-selected) to be 
in a state $\ket \psi$ at time $t = t_1$, all information it is possible to 
know about the system at a subsequent time $t \geq t_1$ is contained in the 
state vector $$\ket{\hat U_{t t_1}\psi}$$ where $\hat U$ denotes its evolution 
in time after measurement at $t_1$, as given by the Schrödinger equation. Using 
this, one can express the quantum amplitude for observing a value $a_m$ of a 
physical observable $A$ at time $t$ as the inner product $$c_m = 
\braket{a_m|\hat U_{t t_1}\psi}$$ with the likelihood of this observation being 
given by ${|c_m|}^2$.

Proponents of TSVF claim that the concept of a quantum state has the limitation 
that it is \textit{time-asymmetric}, as it is defined by measurements made in 
the \textit{past}.  Of course, in classical physics, we are used to measuring a 
system at a time $t_1 < t$ to describe its state at $t$, but this does not 
amount to time-asymmetry. This is because for a classical system, the results 
of measurements performed in the future are already known after measurement 
performed in the past. On the other hand, for a quantum system, measurement 
only partially constrains the possible results of measurements at subsequent 
times, making the standard quantum mechanics formalism truly time-asymmetric.

In Danan et al.'s interferometer, the position of the photon at the beginning 
of the experiment (i.e. at the light source) is known, as well as the position 
at the end (i.e.  at the light detector). That is to say, the system being 
described in the experiment is both pre- and post-selected. The authors claim 
that it is this aspect of the experiment that makes it hard to describe using 
the standard formalism and the associated "common sense" approach to the past 
of a quantum particle.

Consider a simple quantum system that is pre-selected to be in the state
$\ket{\psi_1}$ at time $t_1$ and post-selected to be in the state 
$\ket{\psi_2}$ at time $t_2$. We can consider separately the quantum amplitudes 
of finding it in state $\ket{a_m}$ given that it is in state $\ket{\psi_1}$ at 
$t_1$, $$\braket{a_m|\hat U_{t t_1} \psi_1}$$ and of finding it in the state 
$\bra{\psi_2}$ at $t_2$ given that it is in the state $\ket{a_m}$ at time $t$, 
$$\braket{\psi_2|\hat U_{t t_2} a_m}$$ with $\hat U_{t_a t_b}$ denoting time 
evolution of a state between $t_a$ and $t_b$ as before.

Using standard quantum mechanics, if we want to calculate the likelihood of 
observing the value $a_m$ for the observable $A$ at time $t$ between $t_1$ and 
$t_2$, we can multiply the individual complex amplitudes calculated above, with 
the former representing pre-selection, and the latter representing 
post-selection: $$c_m = \underbrace{\braket{\psi_2|\hat U_{t t_2} 
a_m}}_{post-selection} \underbrace{\braket{a_m|\hat U_{t t_1} 
\psi_1}}_{pre-selection}$$. By performing a mathematical operation under the 
standard quantum mechanical formalism, we can replace $\braket{\psi_2|\hat U_{t 
t_2} a_m}$ with the expression $\braket{\hat U_{t t_2}^{\dagger}\psi_2|a_m}$.  
This gives $$c_m = \braket{\hat U_{t t_2}^\dagger\psi_2| a_m} \braket{a_m|\hat 
U_{t t_1} \psi_1}$$.

We can interpret this as involving the \textit{backwards-evolving} quantum 
state $\bra{\hat U_{t t_2}^{\dagger}\psi_2}$, "prepared" at a time $t_2 > t$.  
TSVF treats the two-state vector $$\bra{\hat U_{t t_2}^{\dagger}\psi_2} 
~\ket{\hat U_{t t_1}\psi}$$ consisting of a backwards-evolving vector and a 
standard forward-evolving vector as containing absolute, concise information 
about a quantum state that has undergone pre and post-selection.

In context of the Tel-Aviv experiment, this means that it is the locations 
where the forwards and backwards-evolving paths coincide that will have their 
frequencies appear in the observed power spectrum. Danan et al. argue that even 
though it is possible to explain their results using the standard quantum 
mechanical formalism\footnote{Using a phenomenon known as 
\textit{super-oscillations}}, it finds its simplest explanation using TSVF, 
where the past and the future are placed on equal footing.

\subsection*{References}

\begin{enumerate}
	\item   Danan, A., Farfurnik, D., Bar-Ad, S., \& Vaidman, L. (2013).  
		Asking Photons Where They Have Been. Phys. Rev. Lett., 
		\textbf{111}, 240402.
	\item  	Aharonov Y., Vaidman L. (2008) The Two-State Vector Formalism: 
		An Updated Review. In: Muga J., Mayato R.S., Egusquiza Í. (eds) 
		Time in Quantum Mechanics.  Lecture Notes in Physics, \textbf{734}.

	\item 	N. Bohr, in Albert Einstein: Philosopher-Scientist, edited by 
		P.  A. Schilpp (Library of Living Philosophers, Evanston, 
		1949); reprinted in Quantum Theory and Measurement, edited by 
		J. A.  Wheeler and W. H. Zurek (Princeton University Press, 
		Princeton, 1983)

	\item	J. S. Lundeen, Physics, \textbf 6, 133 (2013).

\end{enumerate}

\bibliographystyle{plain} % We choose the "plain" reference style
\nocite{*}
\bibliography{refs} % Entries are in the refs.bib file
